Bird beak's elaboration and innovation

### 2023-04-10 Analysis/04-Figure_with_ellipse_scalling_and_8D_histograms.Rmd
Update 04-Figure_with_ellipse_scalling_and_8D_histograms.Rmd

```{r, eval = TRUE, echo = FALSE}
## Get the bar plots for the loadings of the phylogeny component (to be added separately in Inkscape to the main plot)
# pdf(file = "../Manuscript/Figures/phylo_loadings.pdf", height = 4, width = 4)
# plot(NULL, xlim = range(superorder_shapespace$matrix), ylim = range(superorder_shapespace$matrix))
# add.dims(superorder_shapespace, name.col = c(phylogeny = "grey"), n = 1000)
# dev.off()

box_size <- 1.5
pdf(file = "../Manuscript/Figures/ellipses.pdf", height = box_size*10, width = box_size*10)

## SETTING THE WINDOW PARAMETERS
## Global window parameters
lay_mat_super <- rbind(
        matrix(c(1,1, rep(73, 3), 0, 2:5), ncol = 10, nrow = 1), # line 7
        matrix(c(1,1, rep(73, 3), 0, 6:9), ncol = 10, nrow = 1), # line 8
        matrix(c(78,78, 10:17), ncol = 10, nrow = 1)) # line 8)
lay_mat_order <- rbind(
        matrix(c(1,1, rep(57, 3), 60, 2:5), ncol = 10, nrow = 1), 
        matrix(c(1,1, rep(57, 3), 60, 6:9), ncol = 10, nrow = 1), 
        matrix(c(1,1, 10:17), ncol = 10, nrow = 1),
        matrix(c(18:27), ncol = 10, nrow = 1),
        matrix(c(28:37), ncol = 10, nrow = 1), 
        matrix(c(38:47), ncol = 10, nrow = 1),
        matrix(c(48:55, 58:59), ncol = 10, nrow = 1))
heights_vals <- rep(1, 9)
widths_vals <- rep(c(1, 1/3), 5)
layout_matrix <- layout(rbind(lay_mat_super, (lay_mat_order+17)), widths = widths_vals, heights = heights_vals, respect = TRUE)

@@

## Plotting the ellipses
for(i in 1:length(tip_order)) {
    ## Plot each pannel
    wrap.plot.ellipses(i, tip_colours_super_orders, superorder_shapespace, results$super_results_between)
}

## Matching colours (fast utility)
match.col <- function(data, col) {
    return(c(col[match(names(size.subsets(data)), names(col))[-length(size.subsets(data))]], "grey"))
}

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Tinamiformes"= 1,
    "Galliformes"= 3,
    "Anseriformes"= 2,
    "Otidiformes"= 7,
    "Cuculiformes" = 8,
    "Gruiformes" = 9, 
    "Musophagiformes" = 10,
    "Procellariiformes" = 12,
    "Sphenisciformes" = 11,
    "Ciconiiformes" = 13,
    "Pelecaniformes"= 15,
    "Suliformes" = 14,
    "Podicipediformes" = 4,
    "Columbiformes" = 6,
    "Pterocliformes" = 5,
    "Charadriiformes" = 18,
    "Trogoniformes" = 21,
    "Bucerotiformes" = 22,
    "Coraciiformes" = 23,
    "Piciformes" = 24,
    "Strigiformes" = 20,
    "Accipitriformes" = 19, 
    "Falconiformes" = 25,
    "Psittaciformes" = 26,
    "Passeriformes" = 27,
    "Caprimulgiformes" = 16, 
    "Apodiformes" = 17)

@@

## Plotting the ellipses
for(i in 1:length(tip_order)) {
    ## Plot each pannel
    wrap.plot.ellipses(i, tip_colours_orders, order_shapespace, results$order_results_between)
}

## Plotting the aligned and scale ellipses for the superorders
par(bty = "n")
lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = "n", ylab = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
covar.plot(superorder_shapespace, ellipses = mean,
           col = match.col(superorder_shapespace, tip_colours_super_orders),
           legend = FALSE, points = FALSE, scale = "phylogeny",
           lines.lwd = c(rep(1, length(tip_colours_super_orders)), 3),
           add = TRUE, n = 100)

## PLOTTING THE ORDERS
par(bty = "n")
lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = "n", ylab = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
covar.plot(order_shapespace, ellipses = mean,
           col = match.col(order_shapespace, tip_colours_orders),
           legend = FALSE, points = FALSE, scale = "phylogeny",
           lines.lwd = c(rep(0.5, length(tip_colours_orders)), 3),
           add = TRUE, n = 100)
dev.off()
```

>>>>

TODO:
 - [ ] split in two figures
 - [ ] make E/I boxes larger
 - [ ] make E/I colours always orange/blue

```{r, eval = TRUE, echo = FALSE}
## Matching colours (fast utility)
match.col <- function(data, col) {
    return(c(col[match(names(size.subsets(data)), names(col))[-length(size.subsets(data))]], "grey"))
}
```

### Ellipses plot super orders

```{r, eval = TRUE, echo = FALSE}
## Plot dimensions and name
box_size <- 1.5
pdf(file = "../Manuscript/Figures/ellipses_super_order.pdf", height = box_size*4, width = box_size*8)

## New layout for super_order panel (separately)
lay_mat_super_2 <- rbind(
    matrix(c(1,1, rep(18, 3), 0, 2:3), ncol = 8, nrow = 1),
    matrix(c(1,1, rep(18, 3), 0, 4:5), ncol = 8, nrow = 1),
    matrix(c(6:13), ncol = 8, nrow = 1),
    matrix(c(14:17, rep(0, 4)), ncol = 8, nrow = 1))
widths_vals <- rep(c(1, 1/2), 4)
heights_vals <- rep(1, 4)
layout_super <- layout(lay_mat_super_2, widths = widths_vals, heights = heights_vals, respect = TRUE)
# layout.show(layout_super)

@@

## Plotting the ellipses for each order
for(i in 1:length(tip_order)) {
    ## Plot each pannel
    wrap.plot.ellipses(i, tip_colours_super_orders, superorder_shapespace, results$super_results_between, ei.col = c("orange", "blue"))
}

## Adding the ellipses
par(bty = "n")
lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = "n", ylab = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
covar.plot(superorder_shapespace, ellipses = mean,
           col = match.col(superorder_shapespace, tip_colours_super_orders),
           legend = FALSE, points = FALSE, scale = "phylogeny",
           lines.lwd = c(rep(1, length(tip_colours_super_orders)), 3),
           add = TRUE, n = 100)
dev.off()
```

### Ellipses plot orders

```{r, eval = TRUE, echo = FALSE}
## Plot dimensions and name
box_size <- 1.5
pdf(file = "../Manuscript/Figures/ellipses_order.pdf", height = box_size*7, width = box_size*10)

## New layout for order panel (separately)
lay_mat_order_2 <- rbind(
        matrix(c(1,1, rep(56, 3), 0, 2:5), ncol = 10, nrow = 1), 
        matrix(c(1,1, rep(56, 3), 0, 6:9), ncol = 10, nrow = 1), 
        matrix(c(1,1, 10:17), ncol = 10, nrow = 1),
        matrix(c(18:27), ncol = 10, nrow = 1),
        matrix(c(28:37), ncol = 10, nrow = 1), 
        matrix(c(38:47), ncol = 10, nrow = 1),
        matrix(c(48:55, rep(0, 2)), ncol = 10, nrow = 1))
widths_vals <- rep(c(1, 1/2), 5)
heights_vals <- rep(1, 7)
layout_order <- layout(lay_mat_order_2, widths = widths_vals, heights = heights_vals, respect = TRUE)
# layout.show(layout_order)

## Getting the tip order for the ladderised super order tree

## Getting the tip order for the ladderised super order tree
tip_order <- c(
    "Tinamiformes"= 1,
    "Galliformes"= 3,
    "Anseriformes"= 2,
    "Otidiformes"= 7,
    "Cuculiformes" = 8,
    "Gruiformes" = 9, 
    "Musophagiformes" = 10,
    "Procellariiformes" = 12,
    "Sphenisciformes" = 11,
    "Ciconiiformes" = 13,
    "Pelecaniformes"= 15,
    "Suliformes" = 14,
    "Podicipediformes" = 4,
    "Columbiformes" = 6,
    "Pterocliformes" = 5,
    "Charadriiformes" = 18,
    "Trogoniformes" = 21,
    "Bucerotiformes" = 22,
    "Coraciiformes" = 23,
    "Piciformes" = 24,
    "Strigiformes" = 20,
    "Accipitriformes" = 19, 
    "Falconiformes" = 25,
    "Psittaciformes" = 26,
    "Passeriformes" = 27,
    "Caprimulgiformes" = 16, 
    "Apodiformes" = 17)

@@

## Plotting the ellipses
for(i in 1:length(tip_order)) {
    ## Plot each pannel
    wrap.plot.ellipses(i, tip_colours_orders, order_shapespace, results$order_results_between, ei.col = c("orange", "blue"))
}

## Adding the ellipses
par(bty = "n")
lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = "n", ylab = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
covar.plot(order_shapespace, ellipses = mean,
           col = match.col(order_shapespace, tip_colours_orders),
           legend = FALSE, points = FALSE, scale = "phylogeny",
           lines.lwd = c(rep(0.5, length(tip_colours_orders)), 3),
           add = TRUE, n = 100)
dev.off()

```
